---
title: "Linear regression"
author: "Caroline Svahn & Martina Sandberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linear regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



# Description

`linreg` is used to fit linear models.

## Usage

linreg(formula, data)

## Arguments
* formula - an object of class "formula".
* data - an optional data frame.

# Fit a linear model

```{r}
data(iris)
library(RLab4)
model <- linreg(Sepal.Length ~ Petal.Length,iris)
```

`model$plot()` prints out two plots; Residuals vs fitted values and standardized residuals vs fitted values.

```{r}
model$plot()
```


`model$resid()` prints out the vector of residuals.

```{r,results = "hide"}
model$resid()
```

`model$pred()` prints out the predicted values.

```{r,results = "hide"}
model$pred()
```

`model$coef()` prints out the coefficients as a named vector.

```{r}
model$coef()
```

`model$summary()` prints out the coefficients with their standard error, t-value and p-value, the estimeta of $\sigma$ and the degrees of freedom.


```{r}
model$summary()
```

