---
title: "Flight Delay"
author: "Caroline Svahn & Martina Sandberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Flight Delay}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


#Prediction of flight delay

## 1

The data sets `weather` and `flights` are loaded along with the packages `nycflights`, `dplyr`, `caret` and `RLab4` according to:
```{r}
library(nycflights13)
data(flights)
data(weather)
require(dplyr)
library(caret)
library(RLab4)
```


## 2
The datasets are then combined, from `flights` the variable `dep_delay` is saved along with necessary merge variables. From `weather`, all variables except `dewp`, `humid` and `pressure` are used. Two interaction effects are 
included - `prec_wind`, which create effects for preciption and wind speed, and `dir_wind`, creating interation between wind speed and wind direction.

```{r}
flights<- flights %>% select(year, month, day, hour, origin, dep_delay, arr_delay, distance) %>% na.omit()
weather<-weather %>% select(-dewp, -humid, -pressure) %>% na.omit()
wfComb<-flights %>% inner_join(weather, by=c("origin", "year", "month", "day", "hour")) %>% mutate("prec_wind"=precip*wind_speed, "dir_wind"=wind_dir*wind_speed) 

```


## 3
Data partitions are created with
```{r}
trainTestIndex <- createDataPartition(wfComb$dep_delay, p = .85, list = FALSE,times = 1)

Valid  <- wfComb[-trainTestIndex,]
Left <- wfComb[trainTestIndex,]

trainIndex <- createDataPartition(Left$dep_delay, p = 80/85, list = FALSE,times = 1)

Test <- Left[-trainIndex,]
Train <- Left[ trainIndex,] 
```

## 4
Three test models are trained, using `lambda=0.01, 0.001, 0.0001`. Below the code for the best model, with `lambda=0.0001`,is presented. Training to find the best value for lambda confirmed 0.0001 to be beneficial.

```{r}

model7 <- ridgereg(dep_delay ~ distance + temp + wind_dir + wind_speed
                   + precip + visib + prec_wind + dir_wind,
                   data=Train,0.0001)
```


The parameters of the applied model is used to present the RMSE for the validation set:
```{r}
v3<-model7$predict(Valid)
v3Error<-v3-Valid$dep_delay
RMSE3<-sqrt(mean(v3Error^2))

RMSE3
```

## 5
Finally, the RMSE for the Test data is computed.
```{r}
vTest<-model7$predict(Test)
vTestError<-vTest-Test$dep_delay
RMSETest<-sqrt(mean(vTestError^2))

RMSETest
```
which is not far from the RMSE obtained with the training and validation sets, concluding the model seems reliable enough to generalize.

